---
title: "R Notebook"
output: html_notebook
---

# Week 3:

# Part1: Import and export Data

**Check the working directory**

```{r}
getwd()
```

**Set the working directory**

```{r}
setwd("C:/Users/acer/Desktop/Fifth_Semester/Statistical_programming")

```

**read.table() function used to read csv**

```{r}
ABC <- read.table(file= "iris_dirty.csv", header=TRUE, sep= ",")
```

**read.csv() function used to read csv too**

```{r}
ABC <- read.csv("iris_dirty.csv")
```

```{r}
#The result is a data.frame
class(ABC)
```

**help information using**

```{r}
# a help function to find the data directory
file.choose(new=FALSE)
```

```{r}
#to get an explination about the certain function use ?
?read.table
```

**Importing a library**

```{r}
library(readr)
```

**Importing \*.txt File**

```{r}
Mydata1 <- read.table("student_per.txt", sep="\t")

Mydata1
```

```{r}
Mydata2 <- read.delim("student_per.txt")
Mydata2
```

**Importing data from a website**

```{r}
getLink <- "https://calmcode.io/static/data/fish.csv"
Mydata <- read.table(file= getLink, header=TRUE, sep=",")
Mydata
```

**Importing data from excel**

```{r}
library(readxl)
excel_data <- read_excel()
```

**Homework:**

Find the difference between csv, spss, and stata dataset

**Importing data from a package**

```{r}
data(diamonds, package = 'ggplot2')
head(diamonds)
```

## Part 2: Data Cleaning and Pre-processing

**Data Validation steps:**

-   Data loading

-   Getting data summary

-   perform data profiling

```{r}
data <- read.csv("iris_dirty.csv")
data
```

```{r}
summary(data)
```

```{r}
library(DataExplorer)
create_report(data)
```

**Removing Duplicates**

```{r}
iris_new <-unique(data)
rownames(iris_new) <- 1:nrow(iris_new)
sum(duplicated(iris_new))
```

```{r}
cat("This is the old dirty iris dimention: ", dim(data), "\n")

cat("This is the new iris data dimention", dim(iris_new), "\n")

```

**Encoding categorical value**

**factor() or dummyVars():** to encode the values

```{r}
#make sure the data is speelled the same
iris_new$Species <- tolower(iris_new$Species)

#Change to category 
iris_new$Species <- factor(iris_new$Species)

iris_new$Species
```

```{r}
levels(iris_new$Species)[as.integer(iris_new$Species)]
levels(iris_new$Species)[1]<-"setosa"
levels(iris_new$Species)[2]<-"versicolor"
as.numeric(iris_new$Species)
iris_new
```

**Missing data**

**is.na() or complete.cases():** to identify and handle missing values

```{r}
#Handling the misisng data 
#By replacing it with the mean
iris_new$Sepal.Length[is.na(iris_new$Sepal.Length)] <-mean(iris_new$Sepal.Length, na.rm =TRUE)

iris_new$Sepal.Width[is.na(iris_new$Sepal.Width)] <- mean(iris_new$Sepal.Width, na.rm= TRUE)

sum(is.na(iris_new))
```

**Outlier Detection and Treatment**

the use of boxplot, z-score, and mahalanobis distance

```{r}
iris_new <- subset(iris_new, select= -X) #Remove X column

boxplot(iris_new)
```

```{r}
#Deleting outliers
Q1 <- quantile(iris_new$Sepal.Width, .25)
Q3 <- quantile(iris_new$Sepal.Width, .75)

IQR <- IQR(iris_new$Sepal.Width)
```

```{r}
iris_final <- subset(iris, iris_new$Sepal.Width>(Q1-1.5*IQR)&
                       iris_new$Sepal.Width<(Q3 + 1.5*IQR))

boxplot(iris_final)
```
