---
title: "R Notebook"
output: html_notebook
---

# Week 5: Hypothesis test

## **testing one parameter value**

**one sample t-test (parametric)**

```{r}
water<-c(602.4, 604.9, 603.4, 604.0, 600.5, 598.4, 601.8, 600.5, 598.6, 
601.9)
tcalc <- (mean(water)-600)/(sd(water)/sqrt(length(water)))

tcalc

1-pt(tcalc, length(water)-1)
```

```{r}
print("One sided test")
t.test(water, mu=600, alternative= c("greater"), conf.level =0.95)


print("Two sided test")
t.test(water, mu= 600, alternative = c("two.sided"), conf.level = 0.95)

```

### **One tailed test**

1.  **Null Hypothesis (**H0​): The true mean of the water sample is 600.

2.  **Alternative Hypothesis (**H1): The true mean of the water sample is greater than 600.

### Key Values:

-   **t-value**: 2.3921 — This is the calculated t-statistic, which measures how far the sample mean is from the hypothesized mean of 600, in terms of the standard error of the mean.

-   **Degrees of Freedom (df)**: 9 — This refers to the number of independent values in the dataset (sample size - 1).

-   **p-value**: 0.02021 — This is the probability of observing a test statistic as extreme as, or more extreme than, the one calculated (t = 2.3921) assuming the null hypothesis is true.

-   **95% Confidence Interval**: 600.3832 to infinity — This means we are 95% confident that the true population mean is greater than 600.3832.

-   **Sample Mean**: 601.64 — The average value of the water sample.

### Interpretation:

-   Since the p-value (0.02021) is less than the common significance level of 0.05, we reject the null hypothesis. This provides evidence that the true mean of the water sample is **statistically significantly greater than 600** at the 5% significance level.

-   The confidence interval further supports this, showing that the lower bound is 600.3832, indicating that we are 95% confident the true mean is above 600.3832, and there is no upper bound since it is a one-sided test.

### Two Tailed test result

### Interpretation:

-   Since the p-value (0.04042) is **less than 0.05**, we reject the null hypothesis at the 5% significance level. This means there is **statistical evidence that the true mean is significantly different from 600**.

-   The confidence interval [600.0891, 603.1909] does not include 600, further indicating that the true mean is unlikely to be exactly 600. The interval shows that the true mean likely lies between approximately 600.09 and 603.19.

```{r}
#to get the confidience interavel manually 
c(lower = mean(water) - qt(0.975, length(water) - 1)*sd(water)/sqrt(length(water)), Upper = mean(water) + qt(0.975, length(water)- 1)* sd(water)/sqrt(length(water)))
```

## **z-test**

Suppose the IQ in a certain population is normally distributed with a mean of μ = 100 and standard deviation of σ = 15. A scientist wants to know if a new medication affects IQ levels, so she recruits 20 patients to use it for one month and records their IQ levels at the end of the month. Determine if the new medication causes a significant difference in IQ levels.

```{r}
library(BSDA)

cityA = c(82, 84, 85, 89, 91, 91, 92, 94, 99, 99, 105, 109, 109, 109, 110, 112, 112, 113, 114, 114)

z.test(cityA, mu= 100, sigma.x=15) #two tailed

# z.test(cityA, mu= 100, sigma.x=15, conf.level= 0.9, alternative = "greater")

```

The results of the one-sample z-test can be interpreted as follows:

1.  **Null Hypothesis (**H0H_0H0​): The true mean of the population (cityA) is 100.

2.  **Alternative Hypothesis (**H1H_1H1​): The true mean of the population is **not equal** to 100 (two-sided test).

    **Interpretation:**

    -   The **z-value** (0.19379) is very small, indicating that the sample mean (100.65) is close to the hypothesized mean (100).

    -   The **p-value** (0.8463) is much greater than the typical significance level of 0.05. This means we **fail to reject the null hypothesis**, indicating that there is no statistically significant evidence to conclude that the true mean of the population is different from 100.

    -   The **confidence interval** (94.07608 to 107.22392) includes the value 100, which supports the conclusion that the true mean could reasonably be 100.

## Two Population test (Parametric)

### z-test

Suppose the IQ levels among individuals in two different cities are known to be normally distributed each with population standard deviations of 15. A scientist wants to know if the mean IQ level between individuals in city A and city B are different. Determine if the mean IQ level is different between the two cities.

```{r}
cityB = c(90, 91, 91, 91, 95, 95, 99, 99, 108, 109, 109, 114, 115, 116, 117, 117, 128, 129, 130, 133)

z.test(x=cityA, y =cityB, mu=0, sigma.x=15, sigma.y=15) #two sample test 
```

-   The **z-value** of -1.7182 suggests the sample means are somewhat close.

-   The **p-value** of 0.08577 indicates that we do not have enough evidence to reject the null hypothesis at the 5% significance level.

-   The **95% confidence interval** [−17.446925,1.146925] contains 0, meaning that the difference in means could reasonably be 0, further supporting the conclusion.

### t-test

```{r}
x1 <- c(124,120, 121, 118, 119)   # Sample data from population 1
x2 <- c(129, 122, 126, 125, 126, 123)   # Sample data from population 2

# Perform a two-sample t-test
t.test(x1, x2, alternative = c("less"), var.equal = T)

```

-   The **t-value** of -3.2739 indicates a significant difference between the means of x1 and x2.

-   The **p-value** of 0.004812 is less than 0.05, so we reject the null hypothesis.

-   The **confidence interval** (−∞,−2.097692)(-∞, -2.097692)(−∞,−2.097692) supports the alternative hypothesis, showing that the difference in means is significantly negative

## One Population (Non-parametric)

```{r}
wilcox.test(water, mu = 600, alternative = "greater", exact = FALSE)
```

-   **Null Hypothesis (**H0H_0H0​): The true median of the population is equal to 600.

    H0​:median=600

-   **Alternative Hypothesis (**H1H_1H1​): The true median of the population is greater than 600 (one-sided test).

    H1​:median\>600

Based on the p-value of 0.02068, we **reject the null hypothesis** at the 5% significance level. This provides statistical evidence to conclude that the true median is **significantly greater than 600**.

```{r}
# Perform the Wilcoxon signed-rank test
wilcox.test(water, mu = 600, alternative = "two.sided", exact = FALSE)

```

### Two population (Non-parametric )

```{r}
wilcox.test(x1, x2)
wilcox.test(x1, x2, conf.level = 0.9)
wilcox.test(x1, x2, alternative = "less")
```
